\documentclass[border=10pt]{standalone}
%%%<
\usepackage{verbatim}
%%%>
\usepackage{pgfplots}
\pgfplotsset{width=7cm,compat=1.8}
\begin{comment}
:Title: Filling an area with random dots
:Tags: 2D;Random;Filling
:Author: Jake
:Slug: random-fill

Here we fill an area between two curves with random dots.

If the curved had a nearly equal distance, we could
add points between, with some noise generated using a uniform distribution
centered halfway between the curves with a range equal to the offset
between the curves. The higher the number of samples, the denser
the pattern.

However, things get more complicated when you have two curves
that aren't a constant distant apart: in that case, the density of the
random dots would not be constant throughout the domain.

To achieve a constant density, you'll need to adjust the horizontal
distribution of the random points. So instead of the points being a noisy
function of x, we'll use a parametric equation
( x(t), 0.5*a(x(t)) + 0.5*b(x(t)) + U(a(x(t)), b(x(t))).

The horizontal distance between two adjacent points has to be inversely
proportional to the distance between the two bounding curves
(since you need more points to fill a wider band).

To find the horizontal positions for the points that fulfill that
requirements, you can use the antiderivative of the inverse of the difference
of the two bounding curves.

In this case, for example, you would need the antiderivative of
f(x) = 1/(a(x)-b(x)) = 1 / (2*(x/4)^2+1). I think I should be able to do this
by hand, but I cheated and used Wolfram Alpha. The required function is
F(x) = 2 sqrt(2) (atan(x/(2 sqrt(2)))). In order to stretch this function over
the plot domain, you can normalize it using the factor x_max/F(x_max).
The function for the horizontal positions of the points is thus:

F*(t) = 2 sqrt(2) (atan(t/(2*sqrt(2)))) * 5 / 2.99

This code was written by Jake on TeX.SE.
\end{comment}
\begin{document}
\begin{tikzpicture}[
    declare function={a(\x)=(\x/4)^2;},
    declare function={b(\x)=-(\x/4)^2-1;},
    declare function={f(\x) = 2*sqrt(2)*rad(atan(\x/(2*sqrt(2))))*5/2.99;}
]
\begin{axis}[
    domain=0:5, xmin=0,
    axis lines=middle,
    axis equal image,
    xtick=\empty, ytick=\empty,
    enlargelimits=true,
    clip mode=individual, clip=false
]
\addplot [red, only marks, mark=*, samples=500, mark size=0.75]
    (f(x), {0.5*(a(x)+b(x)) + rand * ( a(f(x)) - b(f(x))) / 2});
\addplot [thick] {a(x)};
\addplot [thick] {b(x)};
\end{axis}
\end{tikzpicture}
\end{document}